Training a Gaming Agent on Brainwaves

### 2020-10-18 review2/response.letter.tex

Line 50 “The best overall performance is obtained using Logistic Regression.” // also need to discuss the significance of the overall levels of accuracy in Discussion. Is a best performance of 0.672 useful/acceptable? How does it compare to other researchers (if a comparison is possible).

\begin{quotation}
{\color{blue}
On Section IV, third paragraph we tackled this issue.  The overall performance of 0.672 is low.  Although a straightforward comparison could be misleading in the context of the ErrP experiment, we added a brief recount of the values obtained for other similar works.  We also emphasized this point in future works at the end of the Conclusions section.
}
\end{quotation}

Line 42 col 2 “These results are also consistent with their classification ROC curves, shown \sout{on} $<$in$>$ Figures 6 obtained for both subjects, where the area under the curve are close to chance level.” // Why didn’t you show one good ROC curve and one bad one. This would provide the basis for comparison.

\begin{quotation}
{\color{blue}
Due to previous Reviewer's requests, we added all the ROC curves on the new Figure 6.  We now emphasized on this figure's caption which ROC curves where showing an effective identification of the ErrP potential and which did not.  We added this in the text, highlighting the important finding that we could not provide an enhancement of the agent performance based on the rewards from subjects where the obtained ROC curves were not good (Section III, page 5).
}
\end{quotation}

@@

Page 5
Figure 5 “Y axis shows the averaged number of steps, while x axis show the number of
experiences used to cumulative train the Q-Table.”
// Why does Fig5 E only have 2 sets of data? Some discussion of the (smaller) change from 1 – 2 would be appropriate (e.g. I Discussion)
Line 51 “No\sout{t} performance gain is evidenced, the agents learn nothing which implies that the reward information is useless.” // useless is not a good choice of wording –provides no value

\begin{quotation}
{\color{blue}
We fixed the inappropriate wording, and appreciate the Reviewer for pointing this out.

Figure 5, for Subject 5 (in the first version of the manuscript it was labeled E) has only two sets of testing data because that Participant (OHC) couldn't complete more sessions.  Hence we used two training sessions to calculate the Q-Table and subsequently performed the run session (i.e. the agent performs movements randomly based on the trained QTable for 200 iterations).
On the other hand, we hypothesize that the smaller change obtained from 1-2 for Subject 5 (as well as for Subject 6), may be due to the fact that the ROC curves for those Subjects also display very low classification accuracy.  This could be for reasons related to BCI-illeteracy issues, or the Participant (OHC) not being  concentrated at all during the experiment.  This information was explained in Section III, page 5.
}
\end{quotation}

@@

Line 55 “It can be seen that the overall performance of the agent improves as long as there are more experiences to be used to train it, regardless if they were generated from the brainwaves classification from different subjects.” // more explanation/discussion needed. Is this an average effect from positive learners?

\begin{quotation}
{\color{blue}
This is a great question.  Traditional transfer learning strategies focus on the procedure to enrich a classifier training it with a vast dataset and using it to generalize to new data from unseen scenarios.  While dealing with EEG data, this strategy tends to fail, and other approaches are required.  We found in this work, that even when that strategy fails, if the system is coupled with a basic reinforcement learning algorithm, the cumulative rewards from different subjects, can be used to effectively transmit information and improve the agent performance.  We agree with the Reviewer that this may be due to the average effect of positive or learners with a higher accuracy, but we think it may be related to the fact that the RL algorithm is quite robust and if we can get a higher specificity, the cumulative usage of data from various OHCs help to overcome this issue and improve the agent performance.
}
\end{quotation}

@@

\begin{quotation}
{\color{blue}
The experiment that we performed to generate the Figure 9 doesn't include data from Subject 5 and 6. This is emphasized on the Figure's caption and on the main text in the Result Section.  In the Figure 9, the X axis show the progressive number of gaming agent training matches used to update the Q-Table (in this case they are from different subjects).
}
\end{quotation}

Line 42 “However, even though this implies that the agent misses frequently that an action taken is wrong, this is not hindering the overall performance and the agent is still learning” //

Line 20 col 2 Despite that, the rewards generated from different subjects can be used to train the same Q- Table to improve its performance, which may lead to strategies where the overall performance is improved based on the information from different human critics at the same time.
//These are key findings, worth more discussion/interpretation and inclusion in abstract.

\begin{quotation}
{\color{blue}
We included both finding in the abstract and we emphasized them  in Section IV. 
}
\end{quotation}

@@

Comments to the Author

The paper is improved. The contribution is clearer. Thank you for the reply to comments. I attach some further comments.

The authors have satisfactorily addressed my concerns. I have a minor comment regarding Figure 3. It is unclear what the brown bar labelled GrandAvg represents since there is one for each participant. Please clarify. I suggest having another set of bars (a ninth) for the grand average (average over participants) for each of the algorithms used.

\begin{quotation}
{\color{blue}
We think we misunderstood previous Reviewer's suggestions.  We added the brown bar on the previous Figure 3 (now Figure 4) because we assumed the Reviewer was referring to a classification performed using epoched averaged signal segments.  We used 5-segments to obtain an averaged signal with an enhanced signal-to-noise ratio (Wim van Drongelen, 4 - Signal Averaging, Editor(s): Signal Processing for Neuroscientists, Academic Press, 2007, Pages 55-70,) per Participant, and performed the classification of ErrP components based on those averaged signals for each Participant.  We used only Logistic Regression as classification algorithm, as it was the one with which we achieved better performance for single signal segments.   It is clear now that the brown bars were confusing and we have decided to remove them.  
Instead, we have now included a new Figure 3, where we plotted the \textbf{Grand Average} signals over all the participants for the two experimental conditions.  The top subfigure shows the signals for the "move closer to the target" condition, whereas the bottom subfigure represents the grand average for the "move further" condition.  
On the other hand, Figure 9 shows the results, in terms of the decrease in the number of steps for the agent to reach the goal, of using information to train a classifier with data from one Participant and use it to identify rewards for another participant.  This is performed in a single-trial approach from individual segments without any signal averaging.  
We apologize to the Reviewer for our misunderstanding and ask back for more information if it feels like the misunderstanding persisted.
}
\end{quotation}

### 2020-10-18 ErrP.tex
Pequeños ajustes al leer el texto.

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
%objective
Error-related potential (ErrP) are a particular type of ERP elicited by a person attending a recognizable error.  These Electroencephalographic (EEG)  signals can be used to train a gaming agent by a Reinforcement Learning (RL) algorithm to learn an optimal policy.  The experimental process consists of an observational human critic (OHC) observing a simple game scenario while their brain signals are captured. The game consists of a grid, where a green spot has to reach a desired target in the fewest amount of steps.  Results show that there is an effective transfer of information and that the agent successfully learns to solve the game efficiently, from the initial 97 steps on average required to reach the target to the optimal number of 8 steps.  Our results are expressed in threefold: (i) the mechanics of a simple grid-based game that can elicit the ErrP signal component, (ii) the verification that the reward function only penalizes wrong steps, which means that type II error (not properly identifying a wrong movement) does not affect significantly the agent learning process; (iii) collaborative rewards from multiple observational human critics can be used to train the algorithm effectively and can compensate low classification accuracies (around 67\%) or the limited scope of transfer learning schemes.
%Results
% Expand the results of the paper as this is the contribution. “Results show that there is an effective transfer of information and that the agent learns successfully to solve the game eﬀiciently. Both the underlined terms need to cite evidence from the paper.
% The rewards generated from different subjects can be used to train the same Q- Table to improve its performance, which may lead to strategies where the overall performance is improved based on the information from different human critics at the same time. //These are key findings, worth more discussion/interpretation and inclusion in abstract.
%The algorithm can be trained with information from different OHCs, since the resulting dataset is composed of rewards for each corresponding step, independent of who they were generated from. This shows that the error classification accuracy (approximately 0.67) is good enough for the algorithm to learn. The reward function only penalizes wrong steps, which means that type II error (not properly identifying a wrong movement) does not affect the accuracy, they only make the learning process slower.
%conclusions
%Our results are expressed in threefold (i) the structure of a simple grid-based game that can elicit the ErrP signal component; (ii) the verification that low classification accuracy of just above chance level that produces noisy rewards is enough to allow an agent to learn the optimal policy; (iii) collaborative rewards from multiple observational human critics can compensate the lack of accuracy or the limited scope of transfer learning schemes.

@@

% Reescribí la primera oración porque el sujeto no tenía sentido con el predicado.  TODO
The feasibility of a distinct non-biological communication channel between the Central Nervous System (CNS) and a computer device has been previously proven with Brain Computer Interfaces (BCI) or Brain Machine Interfaces (BMI).~\cite{Vasiljevic2020}.  BCI systems provide a new input modality that can be used in the context of a computer game~\cite{Scherer2012,Nijholt2007}. This advancement is relevant in the context of the accessibility for video games~\cite{Aguado-Delgado2020} and the growing area of e-sports~\cite{Yakovlev2020}.

% Observation human critic? Suena raro acá, puse Subject pero chequear! TODO
In this study, gaming agents are trained using only signal components called Error-related Potentials (ErrP) that can be identified in the observer's brain signals.  These types of signals can be found on EEG traces and are elicited when subjects are aware of the presence of an unexpected outcome, which they identify as an error.  The analysis of ErrP signals is currently an extensive area of research in the neuroscience community~\cite{Holroyd2009}. Error-related Potentials can be detected by signal processing and machine learning techniques~\cite{EERP-PAPER} and are also used in Brain-Computer Interfaces to implement or enhance artificial communication channels~\cite{Chavarriaga2014}.

@@

Previous research has explored the usage of RL with reward signals based on brain activity, recorded by an EEG-based BCI system during task execution. The papers \cite{ROBOT-CONTROL-PAPER,Kim2017,Omedes2013} have successfully demonstrated that a robot can be controlled with brain signals from a person who is observing a robot solve a task.  Moreover, a growing number of studies have demonstrated the feasibility of using ErrPs as rewards for RL schemes such as to enhance robotic behaviour~\cite{Luo2019}, to assess air traffic controller's decisions~\cite{Goh2019} or to categorize actions as errors~\cite{Wirth2020}. Other approaches have used these signals as important feedback for human-robot interaction or to implement shared-control strategies~\cite{Schiatti2018,Iturrate2013}.  Additionally, ErrPs have also been used in the context of games as an additional feedback channel that can be explored to improve gaming experience~\cite{Plass-OudeBos2010,kober2018bci}.

@@

Grand averaged signal segments of 2-seconds length for all the OHCs can be seen on Figure~\ref{fig:grandaverage}.  The ErrP can be noticed more clearly on Pareto-central areas (Pz electrode), around 0.4~s with a more prominent positive peak.

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.5\textwidth]{revisedimages/avg_closer.eps}
    \includegraphics[width=0.5\textwidth]{revisedimages/avg_further.eps}
        \includegraphics[scale=1.0]{revisedimages/electrodes.png}
    \caption{Grand average of 2-seconds segments for all OHCs for the "move closer" (top) and "move further" (middle) condition.  The increased height of the first peak around 0.4 seconds reflects the Error Potential response particularly on Pz, based on the electrode layout (bottom).}
    \label{fig:grandaverage}
\end{figure}

@@

Complementary, Figure \ref{fig:avg_steps} shows the average amount of steps it takes for the agent to reach the goal for each OHC, as the Q-Table is progressively trained using the reward information obtained from the prediction of the trained classifier. Each point corresponds to a run session where the average number of steps for the agent to reach the goal using a specific Q-Table is specified, for 200 repetitions. The first point, at the x-value 0, represents the number of steps the agent takes to reach the goal with an untrained Q-Table, where movements are decided randomly. The next point corresponds to the amount of steps it takes to reach the goal using a policy derived from a Q-Table trained after one brainwave session match, and so on.

@@

Electroencephalographic signals have high inter-subject variability~\cite{Chavarriaga2014}.  This is evidenced in Figure~\ref{fig:transferlearning} where the agent training is performed with rewards obtained by classifying epochs from one Tester OHC and a classifier that was trained using the brainwaves from a different Trainer OHC.  The figure shows the cumulative variation for all run sessions on the average number of steps required to reach the goal after training the agent with all the available matches from the brainwave session.  Enhancements are shown as negative values.  Only the diagonal of the heatmap matrix shows a clear improvement in terms of the reduction of the required number of steps to reach the goal (averaged per 200 runs) which corresponds to the same information for each OHC shown in Figure~\ref{fig:avg_steps}.  For the transfer learning experiment~\cite{Wu2016}, no performance gain is evidenced, the agents learn nothing and this implies that the reward information provides no value.

@@

Each time the gaming agent plays this simple game, it takes on average around 100 steps to reach the target spot.  We trained a classifier to recognize Error Potential from observational human critics that observe the agent playing the game.  The gaming agent plays randomly and movements are marked triggered by the identification of an error potential from the human critic.  Those movements are used as rewards in a Reinforcement Learning scheme to train a Q-Table.  The gaming agent plays the game again, but this time the number of required steps to solve the game is reduced.  If rewards are provided based on random signals,  no reduction is achieved and the average number of steps does not change.   This shows that there is an effective transfer of information from the brainwaves to the agent.   As this process is repeated, the agent keeps improving solving the game effectively, i.e. performing the minimum number of required steps to reach the goal.

@@

This work aims to propose a simple game mechanics that can use the ErrP component to train a gaming agent using a RL model. As described in the literature~\cite{Chavarriaga2014}, ErrPs can be used to transmit subjective feedback to a computer.  The collected data shows that ErrP signals can in fact be classified and used to train an agent effectively.  This proposal tries to keep the system as simple as possible, emphasizing information flow from the subjective error perception of the human critic. Rewards are generated using the signal processing and classification pipeline, and the Q-Table updates, enhancing the performance of the gaming agent.

One additional aspect to remark is the robustness of the learning strategy based on Q-Learning~\cite{Bauer2015,Rubin2012}.   The obtained accuracy to discriminate ErrPs is low, though on the same level to other similar results~\cite{Iturrate2013,Ehrlich2016}.  In this regard, despite the subjectivity of the ErrP response in terms of the proposed experiment, levels of 90\%  accuracy are reported~\cite{Chavarriaga2014} as well.  Despite the low accuracy  in ErrP identification, the RL algorithm was able to extract meaningful information from rewards that were helpful to improve, and often maximize, the agent's performance.  Additionally, one important aspect of the classification results is the low percentage of false positives (Figure~\ref{fig:confusionmatrix}), showing a high specificity. On the other hand, the percentage of false negatives is generally higher.  Even though this implies that the agent misses frequently when a wrong action takes place, this is not hindering the overall performance and the agent is still learning. Although they may be scarce, accurate rewards are very useful for the RL algorithm.

>>>>

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
%objective
Error-related potential (ErrP) are a particular type of Event-Related Potential (ERP) elicited by a person attending a recognizable error.  These Electroencephalographic (EEG)  signals can be used to train a gaming agent by a Reinforcement Learning (RL) algorithm to learn an optimal policy.  The experimental process consists of an observational human critic (OHC) observing a simple game scenario while their brain signals are captured. The game consists of a grid, where a green spot has to reach a desired target in the fewest amount of steps.  Results show that there is an effective transfer of information and that the agent successfully learns to solve the game efficiently, from the initial 97 steps on average required to reach the target to the optimal number of 8 steps.  Our results are expressed in threefold: (i) the mechanics of a simple grid-based game that can elicit the ErrP signal component, (ii) the verification that the reward function only penalizes wrong steps, which means that type II error (not properly identifying a wrong movement) does not affect significantly the agent learning process; (iii) collaborative rewards from multiple observational human critics can be used to train the algorithm effectively and can compensate low classification accuracies and a limited scope of transfer learning schemes.
%Results
% Expand the results of the paper as this is the contribution. “Results show that there is an effective transfer of information and that the agent learns successfully to solve the game eﬀiciently. Both the underlined terms need to cite evidence from the paper.
% The rewards generated from different subjects can be used to train the same Q- Table to improve its performance, which may lead to strategies where the overall performance is improved based on the information from different human critics at the same time. //These are key findings, worth more discussion/interpretation and inclusion in abstract.
%The algorithm can be trained with information from different OHCs, since the resulting dataset is composed of rewards for each corresponding step, independent of who they were generated from. This shows that the error classification accuracy (approximately 0.67) is good enough for the algorithm to learn. The reward function only penalizes wrong steps, which means that type II error (not properly identifying a wrong movement) does not affect the accuracy, they only make the learning process slower.
%conclusions
%Our results are expressed in threefold (i) the structure of a simple grid-based game that can elicit the ErrP signal component; (ii) the verification that low classification accuracy of just above chance level that produces noisy rewards is enough to allow an agent to learn the optimal policy; (iii) collaborative rewards from multiple observational human critics can compensate the lack of accuracy or the limited scope of transfer learning schemes.

@@

% Reescribí la primera oración porque el sujeto no tenía sentido con el predicado.  TODO
The feasibility of a distinct non-biological communication channel between the Central Nervous System (CNS) and a computer device has been previously proven with Brain Computer Interfaces (BCI) or Brain Machine Interfaces (BMI).~\cite{Vasiljevic2020}.  BCI systems provide a new input modality that can be used in computer game~\cite{Scherer2012,Nijholt2007}. This advancement is relevant in the context of the accessibility for video games~\cite{Aguado-Delgado2020} and the growing area of e-sports~\cite{Yakovlev2020}.

% Observation human critic? Suena raro acá, puse Subject pero chequear! TODO
In this study, gaming agents are trained using only signal components called Error-related Potentials (ErrP) that can be identified in the observer's brainwaves.  These types of signals can be found on EEG traces and are elicited when subjects are aware of the presence of an unexpected outcome, which they identify as an error.  The analysis of ErrP signals is currently an extensive area of research in the neuroscience community~\cite{Holroyd2009}. Error-related Potentials can be detected by signal processing and machine learning techniques~\cite{EERP-PAPER} and are also used in Brain-Computer Interfaces to implement or enhance artificial communication channels~\cite{Chavarriaga2014}.

@@

Previous research has explored the usage of RL with reward signals based on brain activity, recorded by an EEG-based BCI system during task execution. The papers \cite{ROBOT-CONTROL-PAPER,Kim2017,Omedes2013} have successfully demonstrated that a robot can be controlled with brain signals from a person who is observing a robot solve a task.  Moreover, a growing number of studies have demonstrated the feasibility of using ErrPs as rewards for RL schemes such as to enhance robotic behaviour~\cite{Luo2019}, to assess air traffic controller's decisions~\cite{Goh2019} or to categorize actions as errors~\cite{Wirth2020}. Other approaches have used these signals as important feedback for human-robot interaction or to implement shared-control strategies~\cite{Schiatti2018,Iturrate2013}.  Additionally, ErrPs have also been used in games as an additional feedback channel that can be explored to improve gaming experience~\cite{Plass-OudeBos2010,kober2018bci}.

@@

Grand averaged time-locked signal segments of 2-seconds length for all the OHCs can be seen in Figure~\ref{fig:grandaverage}.  The ErrP can be noticed more clearly on parieto-central areas (Pz electrode), around 0.4~s with a more prominent positive peak.

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.5\textwidth]{revisedimages/avg_closer.eps}
    \includegraphics[width=0.5\textwidth]{revisedimages/avg_further.eps}
        \includegraphics[scale=1.0]{revisedimages/electrodes.png}
    \caption{Grand average of 2-seconds time-locked segments for all OHCs for the "move closer" (top) and "move further" (middle) condition. Zero corresponds to the onset of the stimulus, i.e. when the gaming agent moves. The increased height of the first peak around 0.4 seconds reflects the ErrP response particularly on Pz, based on the electrode layout (bottom).}
    \label{fig:grandaverage}
\end{figure}

@@

Complementary, Figure \ref{fig:avg_steps} shows the average amount of steps it takes for the agent to reach the goal for each OHC, as the Q-Table is progressively trained using the reward information obtained from the prediction of the trained classifier.  Each point corresponds to a run session of 200 game play repetitions.  The y-value represent the average number of steps the agent takes to reach the goal using a specific Q-Table for a run session. The first point, at the x-value 0, represents the number of steps the agent takes to reach the goal with an untrained Q-Table, where movements are decided randomly. The next point corresponds to the amount of steps it takes to reach the goal using a policy derived from a Q-Table trained after one brainwave session match, and so on.

@@

Electroencephalographic signals have high inter-subject variability~\cite{Chavarriaga2014}.  This is evidenced in Figure~\ref{fig:transferlearning} where the agent training is performed with rewards obtained by classifying epochs from one Tester OHC and a classifier that was trained using the brainwaves from a different Trainer OHC.  The figure shows the cumulative variation for all run sessions on the average number of steps required to reach the goal after training the agent with all the available matches from the brainwave session.  Enhancements are shown as negative values.  Only the diagonal of the heatmap matrix shows a clear improvement in terms of the reduction of the required number of steps to reach the goal (averaged per 200 runs) which corresponds to the same information for each OHC shown in Figure~\ref{fig:avg_steps}.  For this transfer learning experiment~\cite{Wu2016}, no performance gain is evidenced, the agents learn nothing and this implies that the reward information provides no value.

@@

Each time the gaming agent plays this simple game, it takes on average around 100 steps to reach the target spot.  A classifier can be trained to recognize Error Potential from observational human critics that observe the agent playing the game.  The gaming agent plays randomly and movements are marked triggered by the identification of an error potential from the human critic.  Those movements are used as rewards in a Reinforcement Learning scheme to train a Q-Table.  The gaming agent plays the game again, but this time the number of required steps to solve the game is reduced.  If rewards are provided based on random signals,  no reduction is achieved and the average number of steps does not change.   This shows that there is an effective transfer of information from the brainwaves to the agent.   As this process is repeated, the agent keeps improving solving the game effectively, i.e. performing the minimum number of required steps to reach the goal.

@@

This work aims to propose a simple game mechanics that can use the ErrP component to train a gaming agent using a RL model. As described in the literature~\cite{Chavarriaga2014}, ErrPs can be used to transmit subjective feedback to a computer.  The collected data show that ErrP signals can in fact be classified and used to train an agent effectively.  This proposal tries to keep the system as simple as possible, emphasizing information flow from the subjective error perception of the human critic. Rewards are generated using the signal processing and classification pipeline, and the Q-Table updates, enhancing the performance of the gaming agent.

One additional aspect to remark is the robustness of the learning strategy based on Q-Learning~\cite{Bauer2015,Rubin2012}.   The obtained accuracy to discriminate ErrPs is low, though on the same level to other similar results~\cite{Iturrate2013,Ehrlich2016}.  In this regard, despite the subjectivity of the ErrP response in terms of the proposed experiment, levels of 90\%  accuracy are reported~\cite{Chavarriaga2014}.  Despite the low accuracy  in ErrP identification, the RL algorithm was able to extract meaningful information from rewards that were helpful to improve, and often maximize, the agent's performance.  Additionally, one important aspect of the classification results is the low percentage of false positives (Figure~\ref{fig:confusionmatrix}), showing a high specificity. On the other hand, the percentage of false negatives is generally higher.  Even though this implies that the agent misses frequently when a wrong action takes place, this is not hindering the overall performance and the agent is still learning. Although they may be scarce, accurate rewards are very useful for the RL algorithm.
